The galactic chemical evolution of carbon: Constraining stellar nucleosynthesis

### 2023-03-13 carbon_paper/tex/main.tex
better reference names for paper

% Only include extra packages if you really need them. Common packages are:
\usepackage{graphicx}	% Including figure files
\usepackage{amsmath}	% Advanced maths commands
% \usepackage{amssymb}	% Extra maths symbols

@@

%%%%% AUTHORS - PLACE YOUR OWN COMMANDS HERE %%%%%
\defcitealias{2011ApJS..197...17C}{C11}
\defcitealias{2015ApJS..219...40C}{C15}

\newcommand{\cristallo}{\citetalias{2011ApJS..197...17C}+\citetalias{2015ApJS..219...40C}}

@@

\citet{2018MNRAS.476.3432P}
\citep{2017ApJ...837..183W}
\citep{2021MNRAS.508.4484J}
\citep{2021arXiv210603912V}

\citep[e.g.][]{2022arXiv220204666J}

@@

\begin{figure*}
    \centering
    \includegraphics[]{figures/subgiants_mgh.pdf}
    \includegraphics[]{figures/subgiants_mgfe.pdf}
    \caption{Chemical abundance relationships of our sample of APOGEE subgiants (see Jack et al. 2023). Left panel shows }
    \label{fig:sum}
\end{figure*}
\section{Nucleosynthesis}

@@

Taken from J21 and J22, we make the following yield assumptions:
\begin{itemize}
    \item $Y_\text{O}^\text{CC} = 0.015$
    \item $Y_\text{O}^\text{Ia} = 0$
    \item $Y_\text{Fe}^\text{Ia} = 0.00214?$
    \item $Y_\text{Fe}^\text{CC} = 0.0012$
    \item $Y_\text{N}^\text{CC} = 0.00036$
    \item $y_\text{N}^\text{AGB}(M, Z) = 9\times 10^{-4} M \left(\frac{Z}{Z_\odot}\right)$
\end{itemize}
Also following J21/J22, we take the SNeIa Delay Time distribution to be a $t^{-1.1}$ law based on the observations by \citet{2012MNRAS.426.3282M}.

@@

In this work, we explore four different sets of AGB star yields representative of the variety of options which also provide reasonably well sampled grids in metallicity and mass necessary for chemical evolution studies.
\begin{itemize}
    \item~\VICE~\VICE~\VICE
    \item \citetalias{2015ApJS..219...40C}
    \item \cristallo
    \item \cite[C11][Hi]{2011ApJS..197...17C}: \citet{2011ApJS..197...17C, 2015ApJS..219...40C}
    \item K10: \citet{2010MNRAS.403.1413K}
    \item V13: \citet{2013MNRAS.431.3642V}
    \item K16: \citet{2016ApJ...825...26K}
\end{itemize}
Unless otherwise noted, C11 is the model we choose as our fiducial AGB yield.

@@

\begin{figure*}
    \centering
 	    \includegraphics[scale=1]{figures/agb_yields.pdf}\\

@@

    \includegraphics[scale=1]{figures/y_agb_vs_z-1.pdf}
    \includegraphics[scale=1]{figures/y_agb_vs_t-1.pdf}

@@

We summarise many of the predictions of stellar models in Figure \ref{fig:y_cc}. Pink squares represent \citep{2013ARA&A..51..457N}, etc. etc.
\citep{2018ApJS..237...13L}
\citep{2016ApJ...821...38S}
\citep{1995ApJS..101..181W}

On the right, we plot the expected CCSNe abundance ratio of each model as a function of metallicity, defining
\begin{equation}
    {\rm [C/O]^{CC}} = \log_{10}\left( \frac{y_{\rm C}^{\rm CC}}{y_{\rm O}^{\rm CC}}\right) - \log_{10} \left( \frac{Z_{C, \sun }}{Z_{O, \sun }} \right)
\end{equation}
Notice how even these ratios 
\begin{figure*}
    \includegraphics[scale=0.9]{figures/y_c_cc-2.pdf}
    \includegraphics[scale=0.9]{figures/y_co_cc.pdf}
    \caption{The fist plot shows IMF-weighted net fractional CCSNe yield of carbon, $y_C^{cc}$ against metalicity, [M/H], for different studies. The right plot instead shows [C/O]$^\text{CC}$, the logarithm of the abundance ratio of carbon to oxygen if CCSNe were the only carbon source, also as a function of metallicity for the same set of studies. The black line represents the carbon yield of the fiducial model, $y_C^{CC} = 0.004? (Z/Z_{\sun})^{0.3}$, and becomes dashed over regions where we have not tested our model. }
    \label{fig:y_cc}
\end{figure*}

@@

To simulate the chemical evolution of a Milky way like galaxy, we use the Versitile Integrator for Chemical Evolution (\VICE ) published in \citet{2021MNRAS.508.4484J} and also briefly described in \citet{2022arXiv220204666J}. 
As a multi-zone model, the galaxy is divided into 200 rings of 0.1kpc, each with separate gas supplies with outflow and inflow rates.

@@

Because the star formation history is left in a simple inside-out form, this multizone model does not reproduce the high-$\alpha$ sequence, therefore we make the following cut to the V21 data to remove the high alpha sequence. (do we?)

@@

\begin{figure*}
\label{fig:c_evo}
\includegraphics{figures/evo_tracks-9.pdf}
\includegraphics{figures/evo_slices-2.pdf}
\caption{The top set of panes show the time evolution of the gas phase of our fiducial model, plotted as tracks through [C/O]-[O/H] and [C/O]-[O/Fe] where each colored line corresponds to the evolution at that galactic radius. 
The bottom set instead shows the galaxies gas-phase [C/O]-[O/H] and [C/O]-[O/Fe] trend at 5 different time slices. 
}
\end{figure*}

\subsection{Effects of AGB Yields}
\begin{figure*}
\includegraphics[scale=0.6]{figures/oob_agb-1.pdf}

@@

\begin{table}
	\centering
	\caption{For each AGB yield set, our calculation of the effective SFH- and IMF-weighted carbon AGB yield, along with the multiplicative factor to reach an AGB contribution of 20.}
	\label{tab:alpha_agb}
	\begin{tabular}{lcr} % four columns, alignment for each
		\hline
		Study & $y_{\rm C, 0}^{\rm agb}$ & $\alpha_{\rm agb, 20}$\\
		\hline
		C11 & 0.000405 & 3.6\\
		K10 & 0.000643 & 2.3\\
		V13 & 0.000214 & 6.8\\
		K16 & 0.000404 & 3.6\\
		\hline
	\end{tabular}
\end{table}

@@

% \begin{figure*}
% \includegraphics[width=1.5\columnwidth]{figures/cooh_lateburst.pdf}
% \caption{Same as Fig. \ref{fig:stars_abundances} but with different model choices.}
% \end{figure*}

\begin{figure*}
\includegraphics[scale=0.6]{figures/beta-1.pdf}
\includegraphics[scale=0.6]{figures/f_agb-1.pdf}

@@

\begin{figure*}
\includegraphics[scale=0.6]{figures/lateburst_eta-1.pdf}

@@

\begin{figure}
    \includegraphics[scale=0.6]{figures/cooh_scatter.pdf}
    \caption{The simulated stars of our model at present day plotted on [C/O]-[O/H] and colored such that lighter colors represent stars born at greater galactric radii.}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[scale=0.6]{figures/cnoh.pdf}
    \caption{Caption}
    \label{fig:cnofe}
\end{figure}

\begin{figure*}
\centering
\includegraphics[]{figures/summary-2.pdf}
\caption{Present-day gas phase tracks of the fiducial model in [C/O]-[O/H] space. M101 is valid comparison to MW ...(van Dokkum et al. 2014)}
\label{fig:gas_phase}
\end{figure*}
\subsection{The AGB Fraction }
One potential tension with data is our models predict a low value for the AGB fraction of stars.

@@

\begin{figure}
    \centering
    \includegraphics[scale=0.3]{figures/analytic_y-4.pdf}
    \caption{Our reverse fitting method to determine total CCSNe contribution}
\end{figure}

@@

\begin{figure*}
    \centering
    \includegraphics[scale=0.5]{figures/fiorenzo_vs_jack_c-1.pdf}
    \includegraphics[scale=0.5]{figures/fiorenzo_vs_jack_n-1.pdf}
    \caption{A comparison of estimated stellar birth abundance trends from the subgiant sample and \citet{2021arXiv210603912V}
    }
    \label{fig:fiorenzo_jack}
\end{figure*}

@@

@@

\begin{figure*}
\includegraphics[scale=0.45]{figures/agb_adj-2.pdf}
\includegraphics[scale=0.45]{figures/abg_adj_z-1.pdf}

>>>>

% Only include extra packages if you really need them. Common packages are:
\usepackage{graphicx}	% Including figure files
\usepackage{amsmath}	% Advanced maths commands
% \usepackage{amssymb}	% Extra maths symbols
\graphicspath{{../figures/}}

@@

%%%%% AUTHORS - PLACE YOUR OWN COMMANDS HERE %%%%%
\defcitealias{cristallo+11}{C11}
\defcitealias{cristallo+15}{C15}
\defcitealias{ventura+13}{V13}

\newcommand{\cristallo}{\citetalias{cristallo+11}+\citetalias{cristallo+15}}

@@

\citet{prantzos+18}
\citep{WAF17}
\citep{james+21}
\citep{fiorenzo+21}

\citep[e.g.][]{james+22}

@@

\begin{figure*}
    \centering
    \includegraphics[]{subgiants_mgh.pdf}
    \includegraphics[]{subgiants_mgfe.pdf}
    \caption{Chemical abundance relationships of our sample of APOGEE subgiants (see Jack et al. 2023). Left panel shows }
    \label{fig:sum}
\end{figure*}
\section{Nucleosynthesis}

@@

Taken from J21 and J22, we make the following yield assumptions:
\begin{itemize}
    \item $Y_\text{O}^\text{CC} = 0.015$
    \item $Y_\text{O}^\text{Ia} = 0$
    \item $Y_\text{Fe}^\text{Ia} = 0.00214?$
    \item $Y_\text{Fe}^\text{CC} = 0.0012$
    \item $Y_\text{N}^\text{CC} = 0.00036$
    \item $y_\text{N}^\text{AGB}(M, Z) = 9\times 10^{-4} M \left(\frac{Z}{Z_\odot}\right)$
\end{itemize}
Also following J21/J22, we take the SNeIa Delay Time distribution to be a $t^{-1.1}$ law based on
the observations by \citet{maoz+12}.

@@

In this work, we explore four different sets of AGB star yields representative of the variety of options which also provide reasonably well sampled grids in metallicity and mass necessary for chemical evolution studies.
\begin{itemize}
    \item \citetalias{cristallo+11}
    \item \cristallo
    \item \cite[C11][Hi]{cristallo+11}: \citet{cristallo+11}
    \item K10: \citet{karakas10}
    \item V13: \citet{ventura+13}
    \item K16: \citet{KL16}
\end{itemize}
Unless otherwise noted, C11 is the model we choose as our fiducial AGB yield.

@@

\begin{figure*}
    \centering
 	    \includegraphics[scale=1]{agb_yields.pdf}\\

@@

    \includegraphics[scale=1]{y_agb_vs_z-1.pdf}
    \includegraphics[scale=1]{y_agb_vs_t-1.pdf}

@@

We summarise many of the predictions of stellar models in Figure \ref{fig:y_cc}. Pink squares
represent \citep{NKT13}, etc. etc.
\citep{LC18}
\citep{sukhbold+16}
\citep{WW95}

On the right, we plot the expected CCSNe abundance ratio of each model as a function of metallicity, defining
\begin{equation}
    {\rm [C/O]^{CC}} = \log_{10}\left( \frac{y_{\rm C}^{\rm CC}}{y_{\rm O}^{\rm CC}}\right) - \log_{10} \left( \frac{Z_{C, \sun }}{Z_{O, \sun }} \right)
\end{equation}
Notice how even these ratios 
\begin{figure*}
    \includegraphics[scale=0.9]{y_c_cc-2.pdf}
    \includegraphics[scale=0.9]{y_co_cc.pdf}
    \caption{The fist plot shows IMF-weighted net fractional CCSNe yield of carbon, $y_C^{cc}$ against metalicity, [M/H], for different studies. The right plot instead shows [C/O]$^\text{CC}$, the logarithm of the abundance ratio of carbon to oxygen if CCSNe were the only carbon source, also as a function of metallicity for the same set of studies. The black line represents the carbon yield of the fiducial model, $y_C^{CC} = 0.004? (Z/Z_{\sun})^{0.3}$, and becomes dashed over regions where we have not tested our model. }
    \label{fig:y_cc}
\end{figure*}

@@

To simulate the chemical evolution of a Milky way like galaxy, we use the Versitile Integrator for
Chemical Evolution (\VICE\footnote{\VICE~is available at \url{https://github.com/giganano/VICE}}) published in \citet{james+21} and also briefly described in
\citet{james+22}. 
As a multi-zone model, the galaxy is divided into 200 rings of 0.1kpc, each with separate gas supplies with outflow and inflow rates.

@@

Because the star formation history is left in a simple inside-out form, this multizone model does not reproduce the high-$\alpha$ sequence, therefore we make the following cut to the V21 data to remove the high alpha sequence.

@@

\begin{figure*}
\label{fig:c_evo}
\includegraphics{evo_tracks-9.pdf}
\includegraphics{evo_slices-2.pdf}
\caption{The top set of panes show the time evolution of the gas phase of our fiducial model, plotted as tracks through [C/O]-[O/H] and [C/O]-[O/Fe] where each colored line corresponds to the evolution at that galactic radius. 
The bottom set instead shows the galaxies gas-phase [C/O]-[O/H] and [C/O]-[O/Fe] trend at 5 different time slices. 
}
\end{figure*}

\subsection{Effects of AGB Yields}
\begin{figure*}
\includegraphics[scale=0.6]{oob_agb-1.pdf}

@@

@@

% \begin{figure*}
% \includegraphics[width=1.5\columnwidth]{cooh_lateburst.pdf}
% \caption{Same as Fig. \ref{fig:stars_abundances} but with different model choices.}
% \end{figure*}

\begin{figure*}
\includegraphics[scale=0.6]{beta-1.pdf}
\includegraphics[scale=0.6]{f_agb-1.pdf}

@@

\begin{figure*}
\includegraphics[scale=0.6]{lateburst_eta-1.pdf}

@@

\begin{figure}
    \includegraphics[scale=0.6]{cooh_scatter.pdf}
    \caption{The simulated stars of our model at present day plotted on [C/O]-[O/H] and colored such that lighter colors represent stars born at greater galactric radii.}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[scale=0.6]{cnoh.pdf}
    \caption{Caption}
    \label{fig:cnofe}
\end{figure}

\begin{figure*}
\centering
\includegraphics[]{summary-2.pdf}
\caption{Present-day gas phase tracks of the fiducial model in [C/O]-[O/H] space. M101 is valid comparison to MW ...(van Dokkum et al. 2014)}
\label{fig:gas_phase}
\end{figure*}
\subsection{The AGB Fraction }
One potential tension with data is our models predict a low value for the AGB fraction of stars.

@@

\begin{figure}
    \centering
    \includegraphics[scale=0.3]{analytic_y-4.pdf}
    \caption{Our reverse fitting method to determine total CCSNe contribution}
\end{figure}

@@

\begin{figure*}
    \centering
    \includegraphics[scale=0.5]{fiorenzo_vs_jack_c-1.pdf}
    \includegraphics[scale=0.5]{fiorenzo_vs_jack_n-1.pdf}
    \caption{A comparison of estimated stellar birth abundance trends from the subgiant sample and
        \citet{fiorenzo+21}
    }
    \label{fig:fiorenzo_jack}
\end{figure*}

@@

\begin{table}
	\centering
	\caption{For each AGB yield set, our calculation of the effective SFH- and IMF-weighted carbon AGB yield, along with the multiplicative factor to reach an AGB contribution of 20.}
	\label{tab:alpha_agb}
	\begin{tabular}{lcr} % four columns, alignment for each
		\hline
		Study & $y_{\rm C, 0}^{\rm agb}$ & $\alpha_{\rm agb, 20}$\\
		\hline
		C11 & 0.000347 & 2.9\\
		K10 & 0.000585 & 1.7\\
		V13 & 0.000060 & 16.5\\
		K16 & 0.000421 & 2.4\\
		\hline
	\end{tabular}
\end{table}

@@

\begin{figure*}
\includegraphics[scale=0.45]{agb_adj-2.pdf}
\includegraphics[scale=0.45]{abg_adj_z-1.pdf}
